### 2023-05-02 Notes/main.tex
Refactored Simulations/nonmonotone.R into Simulations/generate_data.R and Simulations/nmono_est_funs.R.
Wrote up efficiency comparison with regression estimators. Our methods seem to outperform them.

  \item \textbf{Implemented simulation of nonmonotone MAR data}: 
    Following the approach of \cite{robins1997non}, I construct a nonmonotone
    MAR simulation with two response variables $Y_1$ and $Y_2$ and one covariate
    $X$. The algorithm to generate the data is the following:
    \begin{enumerate}
        \item Generate $X$, $Y_1$, and $Y_2$ for elements $i = 1, \dots, n$.
        \item Using the covariate $X_i$, generate probabilities for each element
          $i$ $p_0$, $p_1$, and $p_2$ such that $p_0 + p_1 + p_2 = 1$. 
        \item Select one option based on the three probabilities for each
          element $i$. If 0 is selected: $R_1 = 0$ and $R_2 = 0$. If 1 is
          selected $R_1 = 1$. If 2 is selected, $R_2 = 1$.
        \item We take the next step in multiple cases. If 0 was selected, we are
          done. If 1 was selected, we generate probabilities $p_{12}$ based on
          $X$ and $Y_1$. Then based on this probability, we determine if $R_2 =
          1$. In the same manner, if 2 was selected in the previous step, we
          generate probabilities $p_{21}$ based on $X$ and $Y_2$. Then based on
          this probability, we determine if $R_2 = 1$.
    \end{enumerate}
    Like the monotone MAR simulation this algorithm produces similar final
    results with the determination of binary variables $R_1$ and $R_2$ and
    variables $X$, $Y_1$, and $Y_2$. Unlike the monotone MAR case, the
    nonmonotone MAR includes observations with $Y_2$ observed and $Y_1$ missing.

@@

    Overall, these results are mostly what I would have expected. All of the
    algorithms estimate the true value of $\theta$ correctly in each case, with
    the oracle estimate having the smallest variance followed by the
    semiparametric algorithm. If there is anything surprising it is that the IPW
    estimator has better performance with the estimated weights compared to the
    true weights. However, I think that this is a known phenomenon.

@@

    \item \textbf{Simulation 2 with Nonmonotone MAR}:
    We also want to simulate data that is correlated. For this simulation, we
    focus on $\Cov(Y_1, Y_2)$. The data generating process now has $\sigma_{yy}
    \neq 0$. We are still interested in $\bar Y_2$ and we still run 2000
    simulation with 2000 observations. In all of the next simulations the true
    value of $\theta = 0$. The results are the following:

@@

    Thus, the proposed method is unbiased with a misspecified outcome model.
    We now show a simulation where the outcome model is correctly specified but
    the response model is not.

@@

    \begin{enumerate}
      \item (Stage 1) Choose a variable observe. We choose $Y_1$ with
        probability $0.4$, $Y_2$ with probability $0.4$ and neither with
        probability $0.2$. If neither, $R_1 = 0$ and $R_2 = 0$. Otherwise,
        continue to Step 2.
      \item (Stage 2) With probability $p_i \propto \logistic(x_i)$, choose to
        observed the other $Y$ variable. 
    \end{enumerate}

@@

  \item However, one of the key steps is that \textit{all} of the response
    models are free of $Y$. In a previous iteration of Simulation 4, we had
    adopted the framework of \cite{robins1997non} where we first to observe the
    first variable and see if we observe the second variable. In this case, the
    second step can depend on the result of the first step and this is what we
    did. However, this makes it the case that $\pi_11$ is a function of $X_i$
    and $Y_1$ and $Y_2$. In this case $\hat \theta_{eff}$ is not unbiased if the
    response model is misspecified (or even just estimated).

@@

  \item To compute the variance, we first solve for each covariance combination.
    Basically, all these computations rely of the following ideas. First, we
    assume that the response model is correctly specified. Consequently,
    $E[A] = E[B] = E[C] = E[D] = 0$ and things work out better. This helps when
    we take the covariance conditional on $X$ because the inner expectations are
    zero. The second key insight it to notice that $E[R_{j}^k] = E[R_j]$ for $j
    \in \{1, 2\}$ and $k \in \mathbb{N}$. This is because $R$ is a binary
    variable. Third, since we assume that the response models are correctly
    specified, we have $E[R_1 \mid X] = \pi_{1+}$, $E[R_2 \mid X] = \pi_{2+}$,
    and $E[R_1, R_2 \mid X] = \pi_{11}$.

    The overall approach to each of this computations is the following:
    (1) take conditional expectations with respect to $X$ (the $\Cov(E[\cdot])$
    term is zero), (2) expand the covariance to $E[XY] - E[X]E[Y]$ (the second
    term is also zero), (3) by the MAR assumption $g, a_2, b_2$ are independent
    of $R_1$ and $R_2$ and we can take the latter out of the expectation, (4)
    evaluate and simplify expressions involving $E[R]$.

@@

\textbf{Questions for Dr. Kim}
\begin{itemize}
  \item Where do we go from here?
  \item This is a functional expectation that I need to calibrate. Should I use
    some sort of basis spline to try to estimate these expectations and then set
    them equal to zero? I don't think that I have worked with this kind of
    constraint before.
\end{itemize}
%\input{prop_full_nonmonoest.tex}

@@

>>>>

  \item \textbf{Implemented simulation of nonmonotone MAR data}: 
    Following the approach of \cite{robins1997non}, I construct a nonmonotone
    MAR simulation with two response variables $Y_1$ and $Y_2$ and one covariate
    $X$. The algorithm to generate the data is the following:
    \begin{enumerate}
        \item Generate $X$, $Y_1$, and $Y_2$ for elements $i = 1, \dots, n$.
        \item Using the covariate $X_i$, generate probabilities for each element
          $i$ $p_0$, $p_1$, and $p_2$ such that $p_0 + p_1 + p_2 = 1$. 
        \item Select one option based on the three probabilities for each
          element $i$. If 0 is selected: $R_1 = 0$ and $R_2 = 0$; if 1 is
          selected $R_1 = 1$; if 2 is selected, $R_2 = 1$.
        \item We take the next step in multiple cases. If 0 was selected, we are
          done. If 1 was selected, we generate probabilities $p_{12}$ based on
          $X$ and $Y_1$. Then based on this probability, we determine if $R_2 =
          1$. In the same manner, if 2 was selected in the previous step, we
          generate probabilities $p_{21}$ based on $X$ and $Y_2$. Then based on
          this probability, we determine if $R_2 = 1$.
    \end{enumerate}
    Like the monotone MAR simulation this algorithm produces similar final
    results with the determination of binary variables $R_1$ and $R_2$ and
    variables $X$, $Y_1$, and $Y_2$. Unlike the monotone MAR case, the
    nonmonotone MAR includes observations with $Y_2$ observed and $Y_1$ missing.

@@

    Overall, these results are mostly what I would have expected. All the
    algorithms estimate the true value of $\theta$ correctly in each case, with
    the oracle estimate having the smallest variance followed by the
    semiparametric algorithm. If there is anything surprising it is that the IPW
    estimator has better performance with the estimated weights compared to the
    true weights. However, I think that this is a known phenomenon.

@@

    \item \textbf{Simulation 2 with Nonmonotone MAR}:
    We also want to simulate data that is correlated. For this simulation, we
    focus on $\Cov(Y_1, Y_2)$. The data generating process now has $\sigma_{yy}
    \neq 0$. We are still interested in $\bar Y_2$ and we still run 2000
    simulation with 2000 observations. In all the next simulations the true
    value of $\theta = 0$. The results are the following:

@@

    Thus, the proposed method is unbiased with a misspecified outcome model.
    We now show a simulation where the outcome model is correctly specified, but
    the response model is not.

@@

    \begin{enumerate}
      \item (Stage 1) Choose a variable observe. We choose $Y_1$ with
        probability $0.4$, $Y_2$ with probability $0.4$ and neither with
        probability $0.2$. If neither, $R_1 = 0$ and $R_2 = 0$. Otherwise,
        continue to Step 2.
      \item (Stage 2) With probability $p_i \propto \logistic(x_i)$, choose to
        observe the other $Y$ variable. 
    \end{enumerate}

@@

  \item However, one of the key steps is that \textit{all} the response
    models are free of $Y$. In a previous iteration of Simulation 4, we had
    adopted the framework of \cite{robins1997non} where we first to observe the
    first variable and see if we observe the second variable. In this case, the
    second step can depend on the result of the first step and this is what we
    did. However, this makes it the case that $\pi_11$ is a function of $X_i$
    and $Y_1$ and $Y_2$. In this case $\hat \theta_{eff}$ is not unbiased if the
    response model is misspecified (or even just estimated).

@@

  \item To compute the variance, we first solve for each covariance combination.
    Basically, all these computations rely on the following ideas. First, we
    assume that the response model is correctly specified. Consequently,
    $E[A] = E[B] = E[C] = E[D] = 0$ and things work out better. This helps when
    we take the covariance conditional on $X$ because the inner expectations are
    zero. The second key insight it to notice that $E[R_{j}^k] = E[R_j]$ for $j
    \in \{1, 2\}$ and $k \in \mathbb{N}$. This is because $R$ is a binary
    variable. Third, since we assume that the response models are correctly
    specified, we have $E[R_1 \mid X] = \pi_{1+}$, $E[R_2 \mid X] = \pi_{2+}$,
    and $E[R_1, R_2 \mid X] = \pi_{11}$.

    The overall approach to each of these computations is the following:
    (1) take conditional expectations with respect to $X$ (the $\Cov(E[\cdot])$
    term is zero), (2) expand the covariance to $E[XY] - E[X]E[Y]$ (the second
    term is also zero), (3) by the MAR assumption $g, a_2, b_2$ are independent
    of $R_1$ and $R_2$ and we can take the latter out of the expectation, (4)
    evaluate and simplify expressions involving $E[R]$.

@@

%\textbf{Questions for Dr. Kim}
%\begin{itemize}
%  \item Where do we go from here?
%  \item This is a functional expectation that I need to calibrate. Should I use
%    some sort of basis spline to try to estimate these expectations and then set
%    them equal to zero? I don't think that I have worked with this kind of
%    constraint before.
%\end{itemize}

%\input{prop_full_nonmonoest.tex}

@@

\input{calibration_comparison.tex}

\input{efficiency_twophase.tex}
